---
title: "Assignment 3"
author: "Sigrid"
date: "2/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

fMRI single voxel data exercise
In this excercise we are going to look at data from an unpublished fMRI experiment. 400 whole-brain EPI
images were acquired for each participant, but in this assignment, we will analyse a time-series from a
single voxel in auditory cortex (transverse temporal gyrus, MNI coordinate: [-46,-20,6], with all time-points
converted into a vector). A total of 37 participants were scanned. They listened to two types of stories
(fiction and factual). A model of the hemodynamic response to the different story types are included in a
separate file. The task is to perform a regresion with these two different independent variables using different
models and also adding an additional covariate.

```{r}
pacman::p_load(tidyverse)
```


```{r cars}
##data
fmri <- as.matrix(read.csv("portfolio_assignment3_aud_fmri_data37.csv", header = FALSE))
##making it a time-series
fmri2 <- ts(fmri)
##design
fmrides <- as.matrix(read.csv("portfolio_assignment3_aud_fmri_design.csv", header=FALSE))
##making it a time-series
fmrides2 <- ts(fmrides)
```

# Task1
Initial figures
1. Make three figures:
1.a. A figure with lineplots of the data from all participants as a function of time in one figure.
1.b. A boxplot with the signal intensity for each participant. Note how much the baseline signal can vary
between participants.
1.c. A lineplots figure with the model covariates.

```{r task 1}

fmri_test <- data.frame(fmri2)
frmi_test <- group_by(ncol)
ggplot(fmri_test, aes(1:400, V1)) + geom_line()

plotmatrix(frmi2)


plot(x = fmri2(1:400,), y = frmi2(,1:37))
plot(fmri2)
vec <- c(fmri2)

plot(vec)

for (i in 1:37) {
  ()
}
  

i <- matrix(rep(1:400), ncol = 1, byrow = T)

fmri_inv <- solve(fmri2)

m <- matrix(c(3,1,1,2), ncol = 2, byrow = T)

fmri_inv <- solve(fmri2)

i <- matrix(c(116,92), ncol = 1)

m_inv %*% i

#first matrix must have the same amount f columns as the second matrix has rows
pacman::p_load(reshape2)
ggplot(melt(fmri2), aes(x = Var1, y = value, color = Var2)) + geom_smooth(method=lm)

#1b
boxplot(fmri2, use.cols = T)
```

