---
title: "Linear algebra"
author: "Mikkel Wallentin"
date: "1/22/2018"
output: pdf_document
---

#Vectors
A vector is basically, a string of numbers, e.g. [1,4,5,6].

Vectors are very useful for a lot of things, and therefore we will spend a bit of time looking at what we can do with vectors.


##Plotting two 2D vectors
```{r }
getwd()
#Two vectors with two elements. 
#We call vectors a 2D vectors.
a1=c(2,2)
a2=c(1,-1)

#Make an empty plot (type='n')
plot(c(-2,3.2),c(-2,3.2),xlim=c(-2.2,4), type="n",xlab='',ylab='',asp=1,frame=FALSE)
#add grid
grid()
#Add arrows for axes
arrows(x0=c(-2, 0),y0=c(0,-2),x1=c(3,0),y1=c(0,3), col=c(1,1))
#Add arrows for vectors
arrows(x0=c(0,0),y0=c(0,0),x1=c(a1[1],a2[1]),y1=c(a1[2],a2[2]),col=c(3,3))
#Add text
text(a1[1]+0.2,a1[2]+0.2,labels='a1=(2,2)')
text(a2[1]-0.2,a2[2]-0.2,labels='a2=(1,-1)')
```

##The length of a 2D vector
We can find the length of a 2D vector using pythagoras' rule for triangles:
a^2+b^2=c^2

```{r}
#A 2D vector
a1=c(2,2)
#Make an empty plot (type='n')
plot(c(-2,3.2),c(-2,3.2),xlim=c(-2.2,4), type="n",xlab='',ylab='',asp=1,frame=FALSE)
#add grid
grid()
#Add arrows for axes
arrows(x0=c(-2, 0),y0=c(0,-2),x1=c(3,0),y1=c(0,3), col=c(1,1))
#Add arrow for vector
arrows(x0=0,y0=0,x1=a1[1],y1=a1[2],col=3)
#Add lines (arrows with no head) to illustrate the triangle
plot1<-arrows(x0=c(0,a1[1]),y0=c(0,0),x1=c(a1[1],a1[1]),y1=c(0,a1[1]),length=0,col=c(2,2))
#Add text
text(a1[1]+0.2,a1[2]+0.2,labels='a1=(x,y)')
text(a1[1]+0.2,a1[2]/2,labels='x')
text(a1[1]/2,+0.2,labels='y')
#text(a1[1]/2-1.2,a1[2]/2+0.2,labels='length(a1=sqrt(x^2+y^2)')

#The actual length
'The length of vector a1 is:'
sqrt(a1[1]^2+a1[2]^2)
```

##Make a plot of a 3D vector
```{r}
library(rgl) # a package for 3D plotting
#Two vectors
b1=c(2,2,2)
b2=c(1,-1,-2)

#Make an empty plot (type='n')
plot3d(c(-2,3.2),c(-2,3.2),c(-2,3.2),
       type="n",xlab='x',ylab='y',zlab='z',asp=1,box=FALSE)

#add grid
grid3d(side=c('x','y+','z'))

#Add arrows for axes
arrow3d(p0=c(-2, 0,0),p1=c(3.2,0,0),col=1,type='lines')
arrow3d(p0=c(0,-2 ,0),p1=c(0,3.2,0),col=1,type='lines')
arrow3d(p0=c(0,0,-2),p1=c(0,0,3.2),col=1,type='lines')

#add vectors
arrow3d(p0=c(0,0,0),p1=b1,col='green',type='lines')
arrow3d(p0=c(0,0,0),p1=b2,col='green',type='lines')

#Add text
texts3d(b1[1]+0.2,b1[2]+0.2,b1[3]+0.2,texts='b1=(2,2,2)')
texts3d(b2[1]-0.2,b2[2]-0.2,b2[3]-0.2,texts='b2=(1,-1,-2)')
```

###Transposing a 15d vector
```{r}
c1=matrix(c(1,2,3,4,5,6,7,8,3,2,3,2,5,6,4),nrow=15)
image(c1,col=rainbow(15))
#transpose t()
c2=t(c1)
image(c2,col=rainbow(15))
```
###Personality Test Example
```{r}
#X=shoe size, Y=balloon_balance,Z=romberg_closed,
morten=c(44,120,120)
sigrid=c(39,3,5)
julie=c(39,3,120)

library(rgl) # a package for 3D plotting

#Make an empty plot (type='n')
plot3d(c(0,100),c(0,100), c(0,120),type="n",
       xlab='shoe size',ylab='balloon_balance',zlab='romberg_closed',asp=1,box=FALSE)

#add grid
grid3d(side=c('x','y+','z'))

#Add arrows for axes
arrow3d(p0=c(0, 0,0),p1=c(100,0,0),col=1,type='lines')
arrow3d(p0=c(0,0 ,0),p1=c(0,100,0),col=1,type='lines')
arrow3d(p0=c(0,0,0),p1=c(0,0,100),col=1,type='lines')

#add vectors
arrow3d(p0=c(0, 0,0),p1=morten,col='green',type='lines')
arrow3d(p0=c(0, 0,0),p1=sigrid,col='green',type='lines')
arrow3d(p0=c(0, 0,0),p1=julie,col='green',type='lines')
#Add text
texts3d(morten[1]+0.2,morten[2]+0.2,morten[3],texts='morten')
texts3d(sigrid[1]-0.2,sigrid[2]-0.2,sigrid[3],texts='sigrid')
texts3d(julie[1]-0.2,julie[2]-0.2,julie[3],texts='julie')
```

###Euclidian distance
```{r}
#add vectors
arrow3d(p0=sigrid,p1=julie,col='red',barblen=0,type='lines')
#Euclidian distance
'The euclidian distance between sigrid and julie is:'
sqrt(sum((sigrid-julie)^2))
```

###Imagine if Julie has a child
```{r}
#Let's make a vector, which is exactly like julie, but half her size
mini_julie=julie/2
#add vectors
arrow3d(p0=c(0,0,0),p1=mini_julie,col='blue',type='lines',lwd=3)
```

###Correlation = small angle in vector space
```{r}
#A correlation between Minijulie and julie
cor.test(mini_julie,julie)
#A correlation between julie and morten
cor.test(morten,julie)
```

###Two vectors are orthogonal if inner produt is 0
```{r}
#Two vectors
a1=c(2,2)
a2=c(1,-1)
#inner product
sum(a1*a2)
```
#Matrices

###Making a matrix
```{r}
#A 2 x 3 matrix
A <- matrix(c(1,3,2,2,8,9),nrow=2,ncol=3)
A
```

###Making out of the personality data
```{r}
#Making matrix out of the personality data
pers <- matrix(c(morten,sigrid,julie),ncol=3)
pers
```

###Transposing the matrix
```{r}
#Transposing the matrix
pers<-t(pers)
pers
```

###The right order can also be achieved by use of byrow()
```{r}
#Making matrix using byrow=TRUE
pers <- matrix(c(morten,sigrid,julie),ncol=3, byrow=TRUE)
pers
```

###Illustrating the matrix using image()
```{r}
#Making matrix using byrow=TRUE
pers <- matrix(c(morten, sigrid,julie),ncol=3,byrow=TRUE)
pers

#convert to image
image(pers,col=gray(1:100/100))
```

Image shows figure at a 90 degree counter-clockwise rotation of the conventional printed layout of a matrix

#Making a function to rotate the image in R
```{r}
#using t() and rev() to rotate the matrix for display
rotate <- function(x) t(apply(x, 2, rev))
pers2<-rotate(pers)
#convert to image
image(pers2,col=gray(1:100/100))
```

#The sum of two matrices
```{r}
#adding two matrices means adding corresponding cells
perspers<-pers + pers
perspers
```

###Multiplying two matrices
```{r}
#If A is an r ?? c matrix and B a c ?? t matrix, i.e. B = [b1 : b2 : ?? ?? ?? : bt]. The product AB is the r ?? t matrix
A<-matrix(1:8,ncol=2)
A
B<-matrix(1:10,nrow=2)
B
#Multiplying two matrices using %*%
C=A%*%B
C
D=B%*%A
D
```

##Some special matrices

??? An n ?? n matrix is a square matrix
??? A matrix A is symmetric if A = A???.
??? A matrix with 0 on all entries is the 0???matrix and is often written simply as 0.
??? A matrix consisting of 1s in all entries is often written J.
??? A square matrix with 0 on all off???diagonal entries and elements d1, d2, . . . , dn on the diagonal a diagonal matrix and is often written diag{d1, d2, . . . , dn}
??? A diagonal matrix with 1s on the diagonal is called the identity matrix and is denoted I. The identity matrix satisfies that IA = AI = A. Likewise, if x is a vector then Ix = x.

###A matrix with ones in all entries
```{r}
#A matrix with ones in all entries
J<-matrix(1,nrow=4,ncol=4)
J
```
###A diagonal matrix
```{r}
#diagonal matrices
A<-diag(c(1:4))
A
```
###A symmetrical matrix
```{r}
#diagonal matrices are symmetrical
A<-diag(c(1:4))
A
t(A)
#A=t(A)
A-t(A)
```
###Identity matrix
```{r}
#A diagonal matrix with 1s on the diagonal is called the identity matrix and is denoted I. The identity matrix satisfies that IA = AI = A. Likewise, if x is a vector then Ix = x.
A<-matrix(c(1:16),nrow=4)
'A:'
A
I<-diag(4)
I
'A*I=A:'
A%*%I
'I*A=A:'
I%*%A
```

###Inverting a matrix
```{r}
#Lets use the personality test data
pers3<-matrix(c(julie, sigrid, morten),nrow=3,byrow=TRUE)
pers3
invpers3<-solve(pers3)
invpers3
#See if we get the identity Matrix by multiplication (and rounding)
round(pers3%*%invpers3)
```

###Let's try with a  matrix with co-linear structure 
```{R}
#What happens if we exchange sigrid with Mini_julie
pers3<-matrix(c(morten, julie, mini_julie),nrow=3,byrow=TRUE)
pers3
solve(pers3)
```

###Solving a linear regression 
```{R}
#Get sleepstudy data from lme4
library(lme4)
#Get participant 308 from sleepstudy
sleep308<-subset(sleepstudy,Subject==308)
attach(sleep308)

#Make a scatter plot
plot(Days,Reaction,type='p')

#Make a dexign matrix with Days and a dummy of ones for getting the intercept
const<-rep(1,10)
X<-matrix(c(Days,const),ncol=2)
X
#Getting the fit
beta<-solve(t(X)%*%X)%*%t(X)%*%Reaction
beta
#Adding a line to the plot Days on X-axis, intercept plus slope*days on y-axis
lines(c(Days[1],Days[10]),c(beta[2]+beta[1]*Days[1],beta[2]+beta[1]*Days[10]),col='darkred')
```

##Extra material

###Indexing vectors and matrices
```{R}
#Use square brackets to index matrices,e.g. 
#the second row of the first column is [2,1]
mat1<-matrix(1:10,ncol=2)
mat1[2,1]
#if you want all rows in a column, you can leave the first index empty
mat1[ ,1]
#and similarly for all columns in a row
mat1[1, ]
```

###Multiplying to matrices element by element
```{R}
#Often you are not interested in matrix multiplication, but only in multiplying the individual elements of two matrices element by element.
#This is easy, the only requirement is that the matrices have the same size

#A 5x2 matrix with numbers 1:10
mat1<-matrix(1:10,ncol=2)
#A 5x2 matrix with 1s in column 1, 2s in column 2
mat2<-matrix(rep(1:2,5),ncol=2,byrow=TRUE)
mat1
mat2
mat1*mat2
```

###Working with vectors and matrices element by element
```{R}
#You can perform addition, subtracttion, multiplication and division
#on vectors and matrices element by element

#A 5x2 matrix with numbers 1:10
mat1<-matrix(1:10,ncol=2)
mat1
(mat1-2)/2

#If they have the same size, you can also take a vector and multiply it on a matrix, e.g. row by row
#Multiply a both columns of a matrix with its own first column
mat1[,1]*mat1
```