---
title: "Portfolio 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
pacman::p_load(tidyverse, reshape2, pastecs, WRS2)
```

```{r importing data}

data <- read.delim("alignment.txt", header = F, sep = "")

```

```{r new dataset - stories}

scan_no <- c(1:400)

onset_story1 <- c(3, 117, 203, 278, 375, 442, 513, 616, 723, 807, 910, 1003, 1093, 1186, 1282) #in seconds

onset_story2 <- c(50, 157, 242, 326, 414, 471, 555, 670, 768, 873, 944, 1054, 1149, 1242, 1316)

onset_ratings <- c(39, 106, 146, 191, 231, 267, 315, 364, 403, 431, 460, 502, 544, 605, 659, 711, 757, 796, 862, 899, 933, 992, 1042, 1082, 1138, 1175, 1231, 1271, 1305, 1348)

duration_story1 <- c(35, 27, 27, 36, 26, 16, 29, 42, 33, 54, 22, 38, 43, 43, 21)

duration_story2 <- c(55, 33, 23, 37, 16, 30, 48, 40, 26, 24, 46, 27, 25, 27, 30)

duration_ratings <- rep(10,10)

ratings_story1 <- c(7,4,4,5,3,1,6,2,1,4,2,3,2,4,1)

ratings_story2 <- c(8,2,3,2,5,7,7,3,3,2,3,3,2,4,3)

df <- data.frame(onset_story1, onset_story2, onset_ratings, duration_story1, duration_story2, duration_ratings, ratings_story1, ratings_story2)

df2 <- data.frame(duration_story1, duration_story2)
df2_melt <- melt(df2)

df2_melt$variable <- as.factor(df2_melt$variable)

```

### Task 1 - Checking input using R

# Test the following hypotheses:
# 1.a. There was a signifiant difference between the durations of the two story types. 

First, I check normality.

```{r checking assumptions + t-test}


round(pastecs::stat.desc(cbind(df$duration_story1, df$duration_story2), basic = F, norm = T), digits = 2)

# or like this??
round(pastecs::stat.desc(df2_melt$value, basic = F, norm = T), digits = 2)

# duration of story 1 could be normally distributed, whereas story 2 is significantly different from a normal distribution.

qqplot_dura_story1 <- ggplot(df, aes(sample = duration_story1))
qqplot_dura_story1 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Duration of story 1 qq-plot")

qqplot_dura_story2 <- ggplot(df, aes(sample = duration_story2))
qqplot_dura_story2 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Duration of story 2 qq-plot")

#doesn't look to shabby

# continuing with both a parametric and a non-parametric t-test

t.test(value ~ variable, df2_melt)
# very high p-value, p = 0.93

#non-parametric t-test
WRS2::yuen(value ~ variable, df2_melt)
#very high p-value 

```

When conducting a non-parametric t-test, there was found no significant difference between the duration of the two stories, t(15.93) = 0.49, p > .05 (p = 0.63).

# 1.b. Hypothesis: There was a significant difference between the ratings of the two story types.

```{r}

df3 <- data.frame(ratings_story1, ratings_story2)
df3 <- melt(df3)

df3$variable <- as.factor(df3$variable)

#checking normality

round(pastecs::stat.desc(cbind(df$ratings_story1, df$ratings_story2), basic = F, norm = T), digits = 2)
#not normally distributed
# or like this
round(pastecs::stat.desc(df3$value, basic = F, norm = T), digits = 2)
# still not normal

qqplot_rat_story1 <- ggplot(df, aes(sample = ratings_story1))
qqplot_rat_story1 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Ratings for story 1 qq-plot")

qqplot_rat_story2 <- ggplot(df, aes(sample = ratings_story2))
qqplot_rat_story2 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Ratings for story 2 qq-plot")
# looks pretty bad

# conducting non-parametric t-test
WRS2::yuen(value ~ variable, df3)


```
When conducting a non-parametric t-test there was found no significant difference betwen the ratings of the two stories, t(15.39) = 0.17, p > .05 (p = 0.87). 

