---
title: "Portfolio 5"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
pacman::p_load(tidyverse, reshape2, pastecs, WRS2)
```

```{r importing data}

data <- read.delim("alignment.txt", header = F, sep = "")

```

```{r new dataset - stories}

#creating the variables
duration_story1 <- c(35, 27, 27, 36, 26, 16, 29, 42, 33, 54, 22, 38, 43, 43, 21)

duration_story2 <- c(55, 33, 23, 37, 16, 30, 48, 40, 26, 24, 46, 27, 25, 27, 30)

ratings_story1 <- c(7,4,4,5,3,1,6,2,1,4,2,3,2,4,1)

ratings_story2 <- c(8,2,3,2,5,7,7,3,3,2,3,3,2,4,3)

# creating two different data frames and melting them for future t-tests

df1 <- data.frame(duration_story1, duration_story2)

df1_melt <- melt(df1)

df1_melt$variable <- as.factor(df1_melt$variable)

df2 <- data.frame(ratings_story1, ratings_story2)

df2_melt <- melt(df2)

df2_melt$variable <- as.factor(df2_melt$variable)

```

# Task 1 - Checking input using R

## Test the following hypotheses:
### 1.a. There was a signifiant difference between the durations of the two story types. 

In order to decide which t-test to conduct the assumption of normality must be checked. If the data are normally distributed, we will conduct a parametric t-test and if not, we'll conduct a non-parametric t-test. 

```{r checking assumptions + t-test}

round(pastecs::stat.desc(cbind(df1$duration_story1, df1$duration_story2), basic = F, norm = T), digits = 2)

# Both variables are not significantly different from a normal distribution according to Shapiro Wilk's test

# looking at the data by making qq-plots
qqplot_dura_story1 <- ggplot(df1, aes(sample = duration_story1))
qqplot_dura_story1 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Duration of story 1 qq-plot")

qqplot_dura_story2 <- ggplot(df1, aes(sample = duration_story2))
qqplot_dura_story2 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Duration of story 2 qq-plot")

# The qq-plots don't look too shabby. We will conclude that the data are normally distributed and we'll continue with conducting a parametric t-test

# parametric t-test

t.test(value ~ variable, df1_melt)

```

When conducting a parametric t-test, there was found no significant difference between the duration of the two stories, t(27.92) = 0.087, p > .05 (p = .93).

### 1.b. Hypothesis: There was a significant difference between the ratings of the two story types.

```{r checking normality + t-test}

#checking normality
round(pastecs::stat.desc(cbind(df2$ratings_story1, df2$ratings_story2), basic = F, norm = T), digits = 2)

# The ratings of story 1 could be normally distributed according to Shapiro Wilk's test, but since the data from the ratings for story 2 are significantly different from a normal distribution, we'll conclude that the data probably are not normally distributed. 

# Visual inspection with qq-plots
qqplot_rat_story1 <- ggplot(df2, aes(sample = ratings_story1))
qqplot_rat_story1 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Ratings for story 1 qq-plot")

qqplot_rat_story2 <- ggplot(df2, aes(sample = ratings_story2))
qqplot_rat_story2 + stat_qq() + stat_qq_line(colour = "red") + ggtitle("Ratings for story 2 qq-plot")

# Data can not be assumed to be normally distributed and we will therefore conduct a non-parametric t-test.

# conducting non-parametric t-test
WRS2::yuen(value ~ variable, df2_melt)


```
When conducting a non-parametric t-test there was found no significant difference betwen the ratings of the two stories, t(15.39) = 0.17, p > .05 (p = .87). 

