---
title: "Assignment 4"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Portfolio assignment 4 - fMRI preprocessing

```{r loading packages}

pacman::p_load(tidyverse, reshape2, Hmisc)

```

# Importing the data from portfolio 3, to be used later in this portfolio. 

```{r port 3}
##data
fmri <- as.matrix(read.csv("portfolio_assignment3_aud_fmri_data37.csv", header = FALSE))
##making it a time-series
fmri2 <- ts(fmri)
##design
fmrides <- as.matrix(read.csv("portfolio_assignment3_aud_fmri_design.csv", header=FALSE))
##making it a time-series
fmrides2 <- ts(fmrides)
```

# Importing the realigned data

```{r loading data}

data <- read.delim("alignment.txt", header = F, sep = "")
```

# Realignment parameters
3.a. Make a lineplot of the realignment parameters in R.

```{r}

data_melt <- melt(data)
data_melt$time <- c(1:400)

ggplot(data_melt, aes(data_melt$time, value, colour = variable)) + geom_line()
```

3.b. How far has the participant moved for each dimension during the experiment (Hint: use “apply()” to run a function across columns)?

```{r}

data_melt %>% group_by(variable) %>% summarise(sum(sqrt(value^2)))

```


3.c. Are any of the realignment parameters significantly correlated with the fMRI model (same model as used in exercise 3)?

```{r}

data_mat <- as.matrix(data, ncol = 6)

cor <- rcorr(data_mat, fmrides2, type = "spearman")
cor

cor_table <- data.frame("Rho_V1" = cor$r[1:6,7], "P-value_V1" = cor$P[1:6,7], "Significant_V1" = ifelse(cor$P[1:6,7] < 0.05, "*", " "),"Rho_V2" = cor$r[1:6,8], "P-value_V2" = cor$P[1:6, 8], "Significant_V2" = ifelse(cor$P[1:6,8]< 0.05, "*", " "))
cor_table
```

Remove linear effects of time from the realignment parameters (hint: 1:400, fit a line and use residuals). 
3.d. Make a lineplot of the realignment parameters with time removed.

```{r}

time <- c(1:400)
model <- lm(data_mat ~ time)
summary(model)

res <- model$residuals

data_melt <- as.matrix(res, ncol= 6)

data_melt <- melt(data_melt)
data_melt$Var2 <- as.factor(data_melt$Var2)
ggplot(data_melt, aes(Var1, value, colour = Var2)) + geom_line() + labs(x = "Picture no.", y = "Value", title = "Realignment parameters with time removed")


```

3.e. Make a lineplot including only the first realignment parameter before and after removal.

```{r}

mat3 <- matrix(c(data[,1], res[,1]), ncol = 2)

mat3_melt <- melt(mat3)
mat3_melt$Var2 <- as.factor(mat3_melt$Var2)

ggplot(mat3_melt, aes(Var1, value, colour = Var2)) + geom_line() + labs(title = "First realignment parameter with original and time corrected data")

# trying something else
# creating a new dataframe 
df_new <- data.frame("Not time corrected" = res[,1], "Time-corrected" = data[,1])

# melting the dataframe (variable/value)
df_melt <- melt(df_new)

# making the dataframe recognize variable 1 (not time corrected) from variable 2 (time corrected)
df_melt$variable <- as.factor(df_melt$variable)

df_melt$picture_no <- c(1:400)

#making a plot
ggplot(df_melt, aes(picture_no, value, colour = variable)) + geom_line() + labs(title = "First realignment parameter with original and time corrected data")

```

3.f. Are the realignment parameters (corrected for effects of time) now correlated with the fMRI model?

```{r}

cor2 <- rcorr(res, fmrides2, type = 'spearman')
cor2

cor_table_2 <- data.frame("Rho_V1" = cor2$r[1:6, 7], "P-value_V1" = cor2$P[1:6, 7], "Significant_V1" = ifelse(cor2$P[1:6,7] < 0.05, "*", " "), "Rho_V2" = cor2$r[1:6, 8], "P-value_V2" = cor2$P[1:6, 8], "Significant_V2" = ifelse(cor2$P[1:6, 8] < 0.05, "*", " "))
cor_table_2

```

